---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  comment = "#>", 
  fig.path = "images/README-"
)
library(gganimate)
```

# SchellingR

The goal of SchellingR is to run the Schelling (1971) agent-based model (ABM) of urban migration and segregation.
The ABM is written in C++ to allow for fast and efficient calculations.
Summary statistics of each round are reported back to R.

## Installation

You can install SchellingR from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("DavZim/SchellingR")
```

## Example

The base function of the package is `run_schelling()` which returns a list of two data.frames `round` and `detailed`.
The former contains aggregated statistics per round, i.e., the number of happy/unhappy agents and the number of moves, the latter contains agent-level information about the grid, the agents, and their neighbors.

The function allows you to tweak basic inputs to the model, such as grid size `size`, the percent of empty grid-cells `percent_empty`, the threshold above which agents are considered to be happy `threshold`, the number of groups to simulated `number_of_groups`, the maximum number of simulation rounds `max_rounds`, and a seed for the random number generator `seed`.

```{r example}
library(SchellingR)
set.seed(1234567)

sh <- run_schelling(size = 10, percent_empty = 0.2, threshold = 0.5, number_of_groups = 2, max_rounds = 100)

str(sh)
```

**For full functionality, you need to have the additional packages installed**

- `gganimate` for animation (gifs)
- `emojifont` for emojis
- `magick` for combining gifs


## Visualisation

After running the simulation, we can plot the average happiness per round using `dplyr` and `ggplot2`.
```{r vis, warning=FALSE, message=F, dependson="example"}
plot_development(sh)
```

We can also use the detailed information to plot the grid at certain steps.
```{r vis2, dependson="example"}
plot_grid(sh, select_round = 8, title = TRUE)
```

## Animation

Using the wonderful [gganimate](https://github.com/thomasp85/gganimate) package, we can also create an animation of the model

```{r vis_animate, warning=FALSE, message=FALSE, cache=TRUE, dependson="example"}
plot_grid(sh, title = TRUE, animate = TRUE)
```

If you want to get even fancier, you can also use the [emojifont](https://CRAN.R-project.org/package=emojifont) package to make the agent's state more intuitive.

```{r vis_emoji, message=FALSE, warning=F, dependson="example"}
plot_grid(sh, select_round = 8, title = TRUE, emoji = TRUE)
```

Or you can use both options at the same time.
```{r vis_anim_emoji, message=FALSE, warning=F, dependson="example"}
plot_grid(sh, title = TRUE, emoji = TRUE, animate = TRUE)
```

## More Complex Example

To give a more complex example on a larger grid with 7 groups:

```{r vis_complex1, dependson="example"}
sh2 <- run_schelling(size = 50, number_of_groups = 5, max_rounds = 1000, seed = 1234567)

grid <- plot_grid(sh2, title = TRUE, animate = TRUE, transition = FALSE)
dev <- plot_development(sh2, animate = TRUE)

combined <- combine_gifs(grid, dev)
print(combined)
```



