---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# SchellingR

The goal of SchellingR is to run the Schelling (1971) agent-based model (ABM) of urban migration and segregation.
The ABM is written in C++ to allow for fast and efficient calculations.
Summary statistics of each round are reported back to R.

## Installation

You can install SchellingR from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("DavZim/SchellingR")
```

## Example

The base function of the package is `run_schelling()` which returns a list of two data.frames `round` and `detailed`.
The former contains aggregated statistics per round, i.e., the number of happy/unhappy agents and the number of moves, the latter contains agent-level information about the grid, the agents, and their neighbors.

The function allows you to tweak basic inputs to the model, such as grid size `size`, the percent of empty grid-cells `percent_empty`, the threshold above which agents are considered to be happy `threshold`, the number of groups to simulated `number_of_groups`, the maximum number of simulation rounds `max_rounds`, and a seed for the random number generator `seed`.

```{r example}
library(SchellingR)
set.seed(1234567)

sh <- run_schelling(size = 10, percent_empty = 0.2, threshold = 0.5, number_of_groups = 2, max_rounds = 100)

str(sh)
```

## Visualisation

After running the simulation, we can plot the average happiness per round using `dplyr` and `ggplot2`.
```{r, warning=F, message=F}
library(dplyr)
library(ggplot2)
theme_set(theme_light())

sh$round %>% 
  mutate(perc_happy = happy_agents / (happy_agents + unhappy_agents)) %>% 
  ggplot(aes(x = round, y = perc_happy)) + 
  geom_line() +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Round", y = "Percent of Agents that are Happy", 
       title = "Happiness Ratio for a Run of the Schelling Model")
```

We can also use the detailed information to plot the grid at certain steps.
```{r}
sh$detailed %>% 
  # group == 0 means empty cells
  filter(group != 0, round == 17) %>% 
  ggplot(aes(x = x, y = y, fill = as.factor(group))) +
  geom_tile() +
  coord_equal() +
  theme_void() +
  scale_fill_brewer(palette = "Set1", guide = FALSE) +
  labs(title = "Schelling Model after 17 Simulation Rounds")
```

## Animation

Using the wonderful [gganimate](https://github.com/thomasp85/gganimate) package, we can also create an animation of the model

```{r, warning=F, message=F}
library(gganimate)

anim <- sh$detailed %>% 
  # group == 0 means empty cells
  filter(group != 0) %>% 
  ggplot(aes(x = x, y = y, fill = as.factor(group), group = id)) +
  geom_tile() +
  coord_equal() +
  theme_void() +
  scale_fill_brewer(palette = "Set1", guide = FALSE) +
  labs(title = "Schelling Model after {previous_state} Simulation Rounds") +
  transition_states(
    round,
    transition_length = 1,
    state_length = 2
  )

animate(anim, nframes = 200)
```

If you want to get even fancier, you can also use the [emojifont](https://CRAN.R-project.org/package=emojifont) package to make the agent's state more intuitive.

```{r, message=F, warning=F}
library(emojifont)

sh$detailed %>% 
  # group == 0 means empty cells
  filter(group != 0, round == 5) %>% 
   mutate(emoji_text = ifelse(happy == 1, emoji("smile"), emoji("angry"))) %>% 
  ggplot(aes(x = x, y = y, fill = as.factor(group), label = emoji_text)) +
  geom_tile() +
  coord_equal() +
  theme_void() +
  scale_fill_brewer(palette = "Set1", guide = FALSE) +
  geom_text(size = 7) +
  labs(title = "Schelling Model after 5 Simulation Rounds")
```


## More Complex Example

To give a more complex example on a larger grid with 7 groups:

```{r}
sh2 <- run_schelling(size = 100, number_of_groups = 7, max_rounds = 10000, seed = 1234567)

sh2$detailed %>% 
  filter(group != 0) %>% 
  group_by(round, group) %>% 
  summarise(perc_happy = mean(happy)) %>% 
  ggplot(aes(x = round, y = perc_happy, color = as.factor(group))) +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Round", y = "Percent Happy", color = "Group",
       title = "Average Happiness per Group")


sh2$detailed %>% 
  filter(group != 0, round == 750) %>%
  ggplot(aes(x = x, y = y, fill = as.factor(group))) + 
  geom_tile() +
  scale_fill_brewer(palette = "Set1") +
  theme_void() +
  labs(title = "More Complex Schelling Model after Round 750", fill = "Group")
```

